AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-05 publications/dora_outlier_detection_iaai.tex
Add description of Earth science dataset. #42

\subsection{Datasets}
Describe the datasets overall and our motivation for choosing these. Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room?

@@

\subsubsection{Targets in Mars rover images.}
Mars exploration is fundamentally an exercise in discovery with the
goal of increasing our understanding of the history, evolution,
composition, and currently active processes on Mars.  Outliers
identified in Mars observations can inspire new discoveries and inform
the choice of which areas merit follow-up, deeper
investigation~\cite{kerner2020comparison,wagstaff:rover-novelty20}.
We collected \todo{XXX} images from the Mars Science Laboratory's
Navigation camera (Navcam) and employed the Rockster~\cite{burl:rockster16}
method used by onboard rover software to identify candidate rock
targets with an area of at least 100 pixels, yielding \num{6005}
targets.  We cropped out a \num{64}$\times$\num{64} pixel image
centered on each target.
%
We collaborated with an MSL science planning team member to
independently review the targets and identify those that were
considered novel by the mission ($n_{novel} = 108$).  Our goal for
this application is to assess how well the selections made by each
algorithm correlate with human novelty judgments to determine which
methods would be most suitable for informing onboard decisions about
follow-up observations.

@@

\section{Acknowledgments}
\todo{Funding/support: JPL CIF, SMD etc.}
We thank the NASA Center Innovation Fund for supporting this work.

\todo{Acknowledge people who contributed to 
datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?}

>>>>

\subsection{Datasets}
We constructed three datasets to evaluate the utility of DORA and the
performance of each ranking algorithm for a variety of scientific domains
 (astrophysics, planetary science, and Earth science). We also included a 
 benchmark dataset that uses MNIST and FashionMNIST. Here we describe
 the details of each dataset.
\todo{Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room? }
 \todo{Can we make any of these datasets publicly available with the paper?}

@@

\subsubsection{Targets in Mars rover images.}
Mars exploration is fundamentally an exercise in discovery with the
goal of increasing our understanding of the history, evolution,
composition, and currently active processes on Mars.  Outliers
identified in Mars observations can inspire new discoveries and inform
the choice of which areas merit follow-up or deeper
investigation~\cite{kerner2020comparison,wagstaff:rover-novelty20}.
We collected \todo{XXX} images from the Navigation camera (Navcam) 
onboard the Mars Science Laboratory (MSL) rover
and employed the Rockster~\cite{burl:rockster16}
method currently used by onboard rover software to identify candidate rock
targets with an area of at least 100 pixels, yielding \num{6005}
targets.  We cropped out a \num{64}$\times$\num{64} pixel image
centered on each target.
%
We collaborated with an MSL science planning team member to
independently review the targets and identify those that were
considered novel by the mission ($n_{novel} = 108$).  Our goal for
this application is to assess how well the selections made by each
algorithm correlate with human novelty judgments to determine which
methods would be most suitable for informing onboard decisions about
follow-up observations.

\subsubsection{Satellite time series for ground-truth observations}
Many Earth science applications using satellite Earth observation (EO) data
 require ground-truth observations for understanding and modeling ground-
 identified objects in the satellite observations. These ground observations
 also serve as labels that are paired with satellite data inputs for machine 
 learning models. For example, a model trained to classify crop types in
 satellite observations requires geo-referenced labels of crop type annotated
 in the field~\cite{tseng2021learning}. A widespread challenge of using ground-
 annotated datasets is that there are often points with erroneous location or 
 label information (e.g., due to GPS location error or human entry error) that
 need to be cleaned before the dataset can be used for machine learning or
 other downstream use cases. Automatically detecting these outliers could save 
 substantial time required for cleaning datasets and improve the performance of
 downstream analyses that rely on high-quality datasets. 
 %
 We used a dataset of ground
 observations of fall army worm (FAW) in maize crops collected by the UN Food
 and Agriculture Organization (FAO). This dataset includes \todo{7XXX} samples
 with location (latitude/longitude) metadata covering \todo{a few countries}. 
 Most locations coincide with crop
 fields but there are many outliers that coincide with other land cover types 
 such as water, buildings, or forests. We constructed an evaluation set 
 consisting of all samples in Kenya ($n=XX$) and manually annotated whether 
 each sample was considered an inlier (i.e., located in a crop field) or not 
 using high-resolution Planet satellite images in Collect Earth 
 Online~\cite{planetdata} ($n_{inlier}=XX$, $n_{outlier}=XX$). We used Google
 Earth Engine to export the Sentinel-1 synthetic aperture radar (SAR) monthly
 time series for each sample location, using data from year each sample was 
 collected. We used SAR data because it is sensitive to ground texture and
  penetrates clouds, which is important for the often-cloudy region covered
  by the dataset. Our goal for this application is to assess how well the 
  selections made by each algorithm correlate with outliers determined by
  visual inspection of the satellite images.
 \todo{If we have time, I want to replace this with a different dataset and 
 experiment that is in progress that evaluates the downstream classifier
 performance before and after applying DORA.}

@@

\section{Acknowledgments}
We thank the NASA Center Innovation Fund and the NASA Science Mission 
Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
 for supporting this work.

\todo{Acknowledge DES?}
We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
(University of Maryland) for providing the fall army worm ground observation
dataset.
